---
title: "R Markdown Baics"
author: "Bradley Rentz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Basic R Markdown is fairly simple to use. The example goes over some of the basic functionality.

Activity 1. The first activity it to try out basic formatting like **bold**, *italics*,superscript^2^ and ~~strikethrough~~. Write some text and try out the basic formating. 

Activity 2. Making bulleted and numbered lists is simple. Add to the example lists below. See how many levels of sub-items you can make.

- List
- Item
    - Sub-item
* Different type of item, same result
  * Sub-item
    
1. Numbered item
1. Second item
    1. Sub-item
1. Third item


Activity 3. R chunks have several options that can be added. Each of the parameters typically takes an option of `FALSE` or `TRUE`, which can be abbreviated as `F` or `T` (capitalization matters).  In the example below change the parameter to `echo=TRUE`. What changes? Try adding the parameter `eval=FALSE` (make sure to add a comma after `echo=FALSE`. What happens? Try `eval=TRUE`. 

```{r unique_name, echo=FALSE}
1 + 3 * 3 * 6^2
```

```{r unique_name2, include=FALSE}
1 + 3 * 3 * 6^2
```

How is `include=FALSE` different from `echo=FALSE`? See if you can find this subtle difference.
Next create a new r code chunk and give it the same name as one of the chunks above, such as `unique_name`. What happens?

Occasionally, when printing an output from R, there are extra text that might be unnecessary. These are typically either `messages` or `warnings`. You can add parameters to the r code chunks to suppress those. In the example below change the `message` and `warning` parameters to `TRUE` one at a time to see what happens.

```{r packages, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
```


Activity 4. Headings. R Markdown allows a several different levels of headings by using `#`. The more `#` that are used, the lower the heading. By default headings are not numbers. To add numbering replace `output: html_document` in the yaml heading at the beginning of the document with:

```{yaml}
output: 
  html_document:
    number_sections: true
```

After changing the numbering, try adding more levels to see what happens. How many levels can be made? The default R Markdown number of numbered levels can be changed. Google how to make that change in the yaml settings above.

# New heading

## Subheading

### Sub-subheading
